### 2022-06-30 doc/manuscript/manuscript.Rmd
Added revisions to the manuscript.rmd file in the pdf file directory

---
title: Identification of high-yielding soybean lines with exceptional seed composition qualities
author:
  - name: Jay Gillenwater
    affil: 1,\ddagger,*
    orcid: 0000-0002-3477-6291
affiliation:
  - num: 1
    address: |
      Department of Crop and Soil Sciences, NCSU, Raleigh, NC, USA
    email: jhgille2@ncsu.edu
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
journal: Agronomy
type: article
status: submit
bibliography: mybibfile.bib
appendix: appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
 In current markets, the primary uses for soybean seed are in products derived from it’s oil or protein content. However, growers are compensated based on seed yield  so a more valuable crop is the one that is high yielding with an optimum combination of protein and oil. A negative correlation of seed protein with seed yield and oil makes simultaneous improvement of  these traits difficult but not impossible through conventional breeding. A selection of lines with exceptional yield and seed composition qualities was made from two recombinant inbred line (RIL) soybean mapping populations to identify high protein and/or high oil lines with yield comparable to elite cultivars.  The performance of these RILs were evaluated in multiple environments and several genotypes were identified with yield comparable to that of high yielding check cultivars with seed protein and/or oil superior to the checks. These genotypes will provide breeders with additional sources of germplasm for continuing efforts to improve seed composition traits without compromising seed yield, and provide growers with more profitable varieties.
keywords: |
  yield; protein; oil;soybean; protein meal
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  The authors declare no conflict of 
  interest.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: RIL
    long: Recombinant inbred line
  - short: RCBD
    long: Randomized complete block design
  - short: CLA
    long: Central crops research station
  - short: CAS
    long: Caswell research farm
  - short: PLY  
    long: Tidewater research station
output: rticles::mdpi_article
---

```{r setup, warning=FALSE,echo = FALSE, message=FALSE}
library(here)
source(here("packages.R"))
tar_load(c(yield_summary_tables, 
           elite_genotype_tables, 
           phenotype_summary_plots))
```

# Introduction

Seed yield, oil, and protein are all valuable traits in a soybean variety, however breeding lines which have both high yield and high protein has been difficult to develop due to the negative correlation between the two traits [@burton1987quantitative; @ProtOilCorr; @ProtOilCorr_new]. While considerable efforts have been made to identify loci which control these seed quality traits so that MAS breeding strategies can be utilized for their improvement, to date the applications of such markers have been few. This is largely due to the lack of markers which are uniquely associated with one trait, and are also stable across genetic and environmental backgrounds. While there is still reason to continue this genetic research, it is important that breeders take every opportunity to identify lines with both high yield, and desirable seed composition traits like oil and protein content so that new value-added varieties can be more profitable.

Soybean lines typically contain about 20% oil and 40% protein content on a dry weight basis [@ProteinGenomics]. The market for soybean meal requires 47.5% protein content in the meal, which corresponds to approximately 41.5% protein content with 21% oil, on a dry weight basis[@ProteinGenomics]. Oil and protein content are two of the most important seed composition traits in soybean, so if one is decreased, the other should be correspondingly increased. The inverse correlation between protein and oil contents is well known and is suspected to be due at least partially to the action of pleiotropic genes and competing metabolic pathways which control the expression of each trait[@SeedCompositionGenomics].

Despite the difficulty in simultaneously breeding for all three of these traits, varieties with elevated protein contents with little or no yield reduction has been released by public sector breeders. The high protein germplasm lines R05-1415 and R05-1772 were released recently that contain 46.9% and 46.1% protein content with 94% and 91% yield, respectively, of the high yielding  cultivar 5002T[@HiProtHiYield02]. Cultivars TN03-350 and TN04-5321 contain 43.9% and 43.1% protein while having superior or comparable yields with checks cultivars[@HiProtHiYield04].). The Highpro1 cultivar was released in 2016 and has a yield which is greater than or equal to that of the check cultivars, while its protein content was 5-6% higher than the checks[@HighPro1].Cultivars produced through conventional breeding techniques such as these have shown that it is possible to identify lines with both high seed protein and seed yield. Efforts to find these lines should continue to provide growers and breeders with additional high value cultivars and germplasm which can be used to further improve protein and yield traits.

To meet this goal, two recombinant inbred line (RIL) mapping populations were screened for lines which showed promising combinations of yield and seed composition traits. Successive rounds of selection were conducted between 2018 and 2021 to identify and characterize lines with high values for yield as well as protein and oil composition.

# Materials and Methods

## Population development

In 2018, oil mapping populations 201 and 202 were grown as plant rows at the Central Crops Research Station in Clayton, NC. These populations consisted of 273 and 237 recombinant inbred lines (RILs) respectively.

The agronomic traits recorded in the field were height, lodging, maturity date, and a composite agronomic score. Lodging was scored on a scale of 1-5 where 5 indicates that all plants in a plot are on the ground, and a score of 1 indicates that all plants are erect[@fehr1987soybeans]. The agronomic score aimed to capture other traits of value such as visual estimation of pod load and plot uniformity to provide a general score of a line’s agronomic desirability. Agronomic score was recorded on a scale of 1-5 as well, with 1 identifying the best lines of a population, and 5 the worst. Maturity was recorded at the R8 maturity date and was recorded as the number of days after September 1. Height was measured in inches from the soil to the top of the plant.

Following harvest, yield, seed weight, protein, and oil content were measured after seed was air dried to approximately 7% moisture content in a greenhouse. Protein and oil contents were measured on a dry basis using a Perten DA 7250 NIR®instrument. Yield and seed weight were measured after seed had been sifted and cleaned of debris.

To select lines for the yield tests in 2019 growing season, lines with low seed yield or extreme maturity dates in 2018 were removed from consideration. Two yield trials were then designed for each mapping population based on maturity dates and desirable agronomic traits, protein and oil contents. Eighty unique lines were selected from each population which satisfied these criteria, and each yield test had 40 RILs. Three high-yielding check cultivars and the two parents of the respective population were also included in each test based on the maturity spread of the RILs. These yield tests were named test 1 and test 2 for RILs derived from mapping population 201 and test 3 and 4 for RILs derived from mapping population 202. Yield check cultivars Dunphy, Osage, and Roy were used in tests 1 and 2, while Dunphy, Dilday, and NC-Raleigh were used for tests 3 and 4. The parents for tests 1 and 2 were cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4 were LMN09-19 and N13-47.

## Experimental design

These four tests were grown in two locations in 2019 - the Tidewater Research Station in Plymouth, NC (PLY) and the Caswell Research Farm in Kinston, NC (CAS). The same data as earlier were collected for each test in this season.

Based the data collected from the 2019 season season, further selections were done to identify fewer lines from the four tests for further yield testing. This was done by identifying the RILs with a yield at or above the average yield of the checks in each test. Further selection was done using the seed composition traits by identifying the thirty RILs with the highest protein + oil content among the RILs which had passed the yield selection threshold.

These thirty lines were then grouped into two new tests of 15 RILs each based on maturity date. These two new tests were named Test 1 and Test 2. Yield check cultivars were again assigned to each test to match the maturity dates of the RILs that were in each test. Cultivars Dunphy, Dilday, and NC-Raleigh were used as checks in Test 1 and Dunphy, Ellis, N10-697, and Osage were used as checks in Test 2.

These two tests were grown in both the 2020 and 2021 seasons. These tests were grown in CLA and CAS in 2020 and CAS and PLY in 2021. RILs were grown in a randomized complete block design with four replications in each location. The same phenotypes were evaluated for each line each season using the same methodology that was employed in the 2019 season.

## Statistical Analysis

Phenotypic traits were analysed with a mixed effects model with the form:

$$y_{ijk} = \mu + E_i + B(E_i) + G_k + GE_{ik} + \epsilon_{ijk}$$

Where $y_{ijk}$ is the phenotypic measurement for rep$j$ of genotype $k$ in environment $i$, $E_i$ is the effect of environment $i$, $B(E_i)$ is the effect of replication nested within environment, $G_k$ is the effect of genotype $G$, $GE_{ik}$ is the interaction effect of environment $E$ and genotype $G$, and $\epsilon_{ijk}$ is the measurement error. The genotype effect was treated as fixed and all other factors were treated as random.

Models were fit using the gamem_met function from the metan package[@olivotoMetanPackageMultienvironment2020]. Least-square means (LS Means) for each genotype and trait were calculated using the above model using the emmeans package[@lenthEmmeansEstimatedMarginal2022] in R. The emmeans package was also used to calculate contrasts as a post-hoc test to compare RIL phenotype means to check means for all collected phenotypes. A sidak adjustment was used to account for multiple comparisons in the calculation of contrasts.Pearson correlation coefficients between each phenotype were calculated with the metan package. Pearson correlation coefficients between each phenotype were calculated with the metan package as well.

The pearson correlation is calculated for each pair of traits as:

$$r = \frac{\sum{(x-m_x)(y-m_y)}}{\sqrt{\sum{(x-m_x)^2}\sum{(y-m_y)^2}}}$$

Where $x$ and $y$ are measurements of the two phenotypes, $m_x$ and $m_y$ are the means of each phenotype, and $r$ is the correlation coefficient.

# Results and discussion

## Phenotypic correlations
A strong negative correlation (r = -0.93) was observed in both tests between seed protein and seed oil content. A weak negative correlation (r = -0.29) was observed between seed protein and yield in both tests which was statistically insignificant in either population. Weak positive correlations were also observed between seed oil and seed yield in both populations. A correlation coefficient of r = 0.24 was observed in Test 1 while a coefficient of r = 0.28 was observed in Test 2. These correlation coefficients were not statistically significant in either population. The pairwise correlation coefficients for Tests 1 and 2 are presented in supplementary figures 2 and 3, respectively.

```{r, lsmean_plot_link, echo = FALSE}

# Get the link to the lsmean scatterplot
mean_plot_link <- phenotype_summary_plots[grepl("lsmean_scatterplot", phenotype_summary_plots)]

```

```{=tex}
\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{`r mean_plot_link`}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}
```

Genotype least square means were calculated for all RILs in both tests. A visualization of these least square means and the standard error on the estimation of these means for seed protein, seed oil, and seed yield can be seen in figure 1. Both tests tended to have more RILs with high protein rather than high oil when compared to the average of the checks, and a moderate number had a yield average greater than the average of the checks. Qualitatively, many RILs had a comparable yield to the checks following the relatively large standard error associated with the estimating the yield, and a qualitatively larger seed protein content given the relatively small standard error associated with the estimation of seed protein content. To separate means quantitatively, we performed contrasts between each RIL mean and the average of the checks for each test.

## Yield contrasts
Several RILs in each of the tests had yield that was comparable to that of their check cultivars as per contrasts between RIL means and the average of the check cultivars for each test. No genotypes had yield that was higher than that of the yield checks, however many had comparable yield. Many of these genotypes with comparable yield also had protein content that was superior to the check average in each test. No genotypes had oil that was superior to the check cultivars, but some had comparable yield and oil content as well as superior protein content. Such genotypes are extremely valuable for production of soybean meal with higher protein. As mentioned earlier, the meal protein of current U.S. cultivars is mostly below the minimum market standard of 47.5%. The meal protein is calculated by a formula that takes both seed protein and oil contents into account (reference from Pantalone). The meal protein contents of these lines with high protein and normal oil contents ranged from xx – xx% compared to the meal protein contents of the checks (Supplementary table XX).

RILs were also evaluated for lodging and seed quality (data not shown?). All RILs had seed quality that was on par with that of the high-yielding checks however, however 11 genotypes had lodging scores that were statistically significantly greater than the checks on the basis of contrasts. These RILs were removed from consideration for recommendation on the basis of their relatively poor agronomic performance

A detailed summary of all contrasts can be seen in supplementary table 1.

## Genotypes with comparable seed yield and seed oil and superior seed protein content
Four genotypes had yield and oil that was similar to that of the yield checks, and seed protein that was greater than the average of the checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627, N18-1643, and N18-1783 from Test 2.  Summary data for these lines is given in table 1.

```{r elite_geno_table, results = "asis", echo = FALSE}

print(elite_genotype_tables$elite_genos)

```

A strong negative correlation was observed in both populations between seed oil and seed protein, and a weak inverse correlation was observed between seed protein and seed yield. This observation matches well with previous findings on the inverse correlation between seed protein and seed oil, and between seed protein and seed yield. Many RILs had high protein when compared with the check cultivars so it was not surprising that relatively few had both a superior seed protein content, and a comparable seed oil content.

These RILs can provide valuable germplasm that can be used to simultaneously improve both seed oil and seed protein content without compromising seed yield or agronomics. This can be of use both to breeders who are looking to improve seed composition traits in soybean and for growers who are seeking options for novel soybean varieties with seed composition that is superior to that of elite cultivars.

## Genotypes with comparable yield and superior protein
Seventeen RILs had a similar yield and superior protein content to the check cultivars. Summary data for the seed protein and seed yield of these lines is given in table 2. From among these RILs, N18-1627 had the highest seed yield relative to the checks included in the test with an average yield that was 105.9% that of the check average of Test 2. Genotype N18-1763 had the highest seed protein content on average with a protein content that was 116.8% that of the check average of Test 1. We observed a negative correlation between seed protein and seed yield in both tests so RILs with seed protein content that was above the check average tended to have a seed yield content that was comparatively lower.

```{r high_po_table, results = "asis", echo = FALSE}

print(elite_genotype_tables$hp_sy_table)

```

However, among these genotypes are several with both average seed protein and average seed yield that were greater than the averages of the checks for each test. These N18-1632-1, N18-1620, N18-1635 from test 1 and RILs N18-1761 and N18-1627 from test 2. Of particular note are genotypes N18-1632-1 from test 1 and N18-1761 from test 2. These two RILs have a seed protein content that is substantially above that of the average protein content of he checks while also maintaining an average yield that is above the average yield of the checks. These RILs are ideal candidates for use in breeding programs which seek to improve seed protein content in particular without compromising seed yield or argonomics.

# Conclusion

We have identified several genotypes with seed yield that is comparable to existing cultivars that are commonly used in production. Many of these genotypes have seed protein content that exceeds that of the check cultivars, and some of these with superior protein have a seed oil content that is comparable to these check cultivars as well. These genotypes have good agronomic qualities as well and will provide both breeders and growers with new options for genotypes that can be used in production or used in breeding programs that seek to improve valuable soybean seed composition traits with significantly reducing the yield.