### 2022-06-30 doc/manuscript/manuscript.tex
Added a a step to calculate protein meal and re-ran the pipeline. lsmeans seem to have been calculated without issue.
Also added protein meal to the phenotype name conversion utility tables.

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[Agronomy,article,submit,moreauthors,pdftex]{mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%% Some pieces required from the pandoc template
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2019}
\copyrightyear{2019}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Identification of high-yielding soybean lines with exceptional
seed composition qualities}

% Authors, for the paper (add full first names)
\Author{Jay
Gillenwater$^{1,\ddagger,*}$\href{https://orcid.org/0000-0002-3477-6291}{\orcidicon}}

% Authors, for metadata in PDF
\AuthorNames{Jay Gillenwater}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Department of Crop and Soil Sciences, NCSU, Raleigh, NC,
USA; \href{mailto:jhgille2@ncsu.edu}{\nolinkurl{jhgille2@ncsu.edu}}\\
}
% Contact information of the corresponding author
\corres{Correspondence: }

% Current address and/or shared authorship
\firstnote{Current address: Updated affiliation}
\secondnote{These authors contributed equally to this work.}

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
\simplesumm{A Simple summary goes here.}

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{In current markets, the primary uses for soybean seed are in
products derived from it's oil or protein content. However, growers are
compensated based on seed yield so a more valuable crop is the one that
is high yielding with an optimum combination of protein and oil. A
negative correlation of seed protein with seed yield and oil makes
simultaneous improvement of these traits difficult but not impossible
through conventional breeding. A selection of lines with exceptional
yield and seed composition qualities was made from two recombinant
inbred line (RIL) soybean mapping populations to identify high protein
and/or high oil lines with yield comparable to elite cultivars. The
performance of these RILs were evaluated in multiple environments and
several genotypes were identified with yield comparable to that of high
yielding check cultivars with seed protein and/or oil superior to the
checks. These genotypes will provide breeders with additional sources of
germplasm for continuing efforts to improve seed composition traits
without compromising seed yield, and provide growers with more
profitable varieties.}

% Keywords
\keyword{yield; protein; oil;soybean; protein meal}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Seed yield, oil, and protein are all valuable traits in a soybean
variety, however breeding lines which have both high yield and high
protein has been difficult to develop due to the negative correlation
between the two traits
\citep{burton1987quantitative, ProtOilCorr, ProtOilCorr_new}. While
considerable efforts have been made to identify loci which control these
seed quality traits so that MAS breeding strategies can be utilized for
their improvement, to date the applications of such markers have been
few. This is largely due to the lack of markers which are uniquely
associated with one trait, and are also stable across genetic and
environmental backgrounds. While there is still reason to continue this
genetic research, it is important that breeders take every opportunity
to identify lines with both high yield, and desirable seed composition
traits like oil and protein content so that new value-added varieties
can be more profitable.

Soybean lines typically contain about 20\% oil and 40\% protein content
on a dry weight basis \citep{ProteinGenomics}. The market for soybean
meal requires 47.5\% protein content in the meal, which corresponds to
approximately 41.5\% protein content with 21\% oil, on a dry weight
basis\citep{ProteinGenomics}. Oil and protein content are two of the
most important seed composition traits in soybean, so if one is
decreased, the other should be correspondingly increased. The inverse
correlation between protein and oil contents is well known and is
suspected to be due at least partially to the action of pleiotropic
genes and competing metabolic pathways which control the expression of
each trait\citep{SeedCompositionGenomics}.

Despite the difficulty in simultaneously breeding for all three of these
traits, varieties with elevated protein contents with little or no yield
reduction has been released by public sector breeders. The high protein
germplasm lines R05-1415 and R05-1772 were released recently that
contain 46.9\% and 46.1\% protein content with 94\% and 91\% yield,
respectively, of the high yielding cultivar
5002T\citep{HiProtHiYield02}. Cultivars TN03-350 and TN04-5321 contain
43.9\% and 43.1\% protein while having superior or comparable yields
with checks cultivars\citep{HiProtHiYield04}.). The Highpro1 cultivar
was released in 2016 and has a yield which is greater than or equal to
that of the check cultivars, while its protein content was 5-6\% higher
than the checks\citep{HighPro1}.Cultivars produced through conventional
breeding techniques such as these have shown that it is possible to
identify lines with both high seed protein and seed yield. Efforts to
find these lines should continue to provide growers and breeders with
additional high value cultivars and germplasm which can be used to
further improve protein and yield traits.

To meet this goal, two recombinant inbred line (RIL) mapping populations
were screened for lines which showed promising combinations of yield and
seed composition traits. Successive rounds of selection were conducted
between 2018 and 2021 to identify and characterize lines with high
values for yield as well as protein and oil composition.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{population-development}{%
\subsection{Population development}\label{population-development}}

In 2018, oil mapping populations 201 and 202 were grown as plant rows at
the Central Crops Research Station in Clayton, NC. These populations
consisted of 273 and 237 recombinant inbred lines (RILs) respectively.

The agronomic traits recorded in the field were height, lodging,
maturity date, and a composite agronomic score. Lodging was scored on a
scale of 1-5 where 5 indicates that all plants in a plot are on the
ground, and a score of 1 indicates that all plants are
erect\citep{fehr1987soybeans}. The agronomic score aimed to capture
other traits of value such as visual estimation of pod load and plot
uniformity to provide a general score of a line's agronomic
desirability. Agronomic score was recorded on a scale of 1-5 as well,
with 1 identifying the best lines of a population, and 5 the worst.
Maturity was recorded at the R8 maturity date and was recorded as the
number of days after September 1. Height was measured in inches from the
soil to the top of the plant.

Following harvest, yield, seed weight, protein, and oil content were
measured after seed was air dried to approximately 7\% moisture content
in a greenhouse. Protein and oil contents were measured on a dry basis
using a Perten DA 7250 NIR®instrument. Yield and seed weight were
measured after seed had been sifted and cleaned of debris.

To select lines for the yield tests in 2019 growing season, lines with
low seed yield or extreme maturity dates in 2018 were removed from
consideration. Two yield trials were then designed for each mapping
population based on maturity dates and desirable agronomic traits,
protein and oil contents. Eighty unique lines were selected from each
population which satisfied these criteria, and each yield test had 40
RILs. Three high-yielding check cultivars and the two parents of the
respective population were also included in each test based on the
maturity spread of the RILs. These yield tests were named test 1 and
test 2 for RILs derived from mapping population 201 and test 3 and 4 for
RILs derived from mapping population 202. Yield check cultivars Dunphy,
Osage, and Roy were used in tests 1 and 2, while Dunphy, Dilday, and
NC-Raleigh were used for tests 3 and 4. The parents for tests 1 and 2
were cultivars LMN09-119 and N09-09, and the parents for tests 3 and 4
were LMN09-19 and N13-47.

\hypertarget{experimental-design}{%
\subsection{Experimental design}\label{experimental-design}}

These four tests were grown in two locations in 2019 - the Tidewater
Research Station in Plymouth, NC (PLY) and the Caswell Research Farm in
Kinston, NC (CAS). The same data as earlier were collected for each test
in this season.

Based the data collected from the 2019 season season, further selections
were done to identify fewer lines from the four tests for further yield
testing. This was done by identifying the RILs with a yield at or above
the average yield of the checks in each test. Further selection was done
using the seed composition traits by identifying the thirty RILs with
the highest protein + oil content among the RILs which had passed the
yield selection threshold.

These thirty lines were then grouped into two new tests of 15 RILs each
based on maturity date. These two new tests were named Test 1 and Test
2. Yield check cultivars were again assigned to each test to match the
maturity dates of the RILs that were in each test. Cultivars Dunphy,
Dilday, and NC-Raleigh were used as checks in Test 1 and Dunphy, Ellis,
N10-697, and Osage were used as checks in Test 2.

These two tests were grown in both the 2020 and 2021 seasons. These
tests were grown in CLA and CAS in 2020 and CAS and PLY in 2021. RILs
were grown in a randomized complete block design with four replications
in each location. The same phenotypes were evaluated for each line each
season using the same methodology that was employed in the 2019 season.

\hypertarget{statistical-analysis}{%
\subsection{Statistical Analysis}\label{statistical-analysis}}

Phenotypic traits were analysed with a mixed effects model with the
form:

\[y_{ijk} = \mu + E_i + B(E_i) + G_k + GE_{ik} + \epsilon_{ijk}\]

Where \(y_{ijk}\) is the phenotypic measurement for rep\(j\) of genotype
\(k\) in environment \(i\), \(E_i\) is the effect of environment \(i\),
\(B(E_i)\) is the effect of replication nested within environment,
\(G_k\) is the effect of genotype \(G\), \(GE_{ik}\) is the interaction
effect of environment \(E\) and genotype \(G\), and \(\epsilon_{ijk}\)
is the measurement error. The genotype effect was treated as fixed and
all other factors were treated as random.

Models were fit using the gamem\_met function from the metan
package\citep{olivotoMetanPackageMultienvironment2020}. Least-square
means (LS Means) for each genotype and trait were calculated using the
above model using the emmeans
package\citep{lenthEmmeansEstimatedMarginal2022} in R. The emmeans
package was also used to calculate contrasts as a post-hoc test to
compare RIL phenotype means to check means for all collected phenotypes.
A sidak adjustment was used to account for multiple comparisons in the
calculation of contrasts.Pearson correlation coefficients between each
phenotype were calculated with the metan package. Pearson correlation
coefficients between each phenotype were calculated with the metan
package as well.

The pearson correlation is calculated for each pair of traits as:

\[r = \frac{\sum{(x-m_x)(y-m_y)}}{\sqrt{\sum{(x-m_x)^2}\sum{(y-m_y)^2}}}\]

Where \(x\) and \(y\) are measurements of the two phenotypes, \(m_x\)
and \(m_y\) are the means of each phenotype, and \(r\) is the
correlation coefficient.

\hypertarget{results-and-discussion}{%
\section{Results and discussion}\label{results-and-discussion}}

\hypertarget{phenotypic-correlations}{%
\subsection{Phenotypic correlations}\label{phenotypic-correlations}}

A strong negative correlation (r = -0.93) was observed in both tests
between seed protein and seed oil content. A weak negative correlation
(r = -0.29) was observed between seed protein and yield in both tests
which was statistically insignificant in either population. Weak
positive correlations were also observed between seed oil and seed yield
in both populations. A correlation coefficient of r = 0.24 was observed
in Test 1 while a coefficient of r = 0.28 was observed in Test 2. These
correlation coefficients were not statistically significant in either
population. The pairwise correlation coefficients for Tests 1 and 2 are
presented in supplementary figures 2 and 3, respectively.

\begin{figure}[H]
\centering
\includegraphics[width=15 cm]{C:/Users/jhgille2/Documents/R/projects/Jay_yield/exports/plots/lsmean_scatterplot.pdf}
\caption{Genotype Least square means for seed oil, seed protein, and seed yield for soybean RILs in Test 1 (\textbf{A}) and Test 2 (\textbf{B}). Points indicate the least square means and error bars indicate the standard errors in the estimation of least square means. Blue dots indicate that a RIL has a least square mean above the average of the checks in the test and a color of red indicates that the RIL has a mean value below the check average. The check average is shown with the vertical dashed line.}
\end{figure}

Genotype least square means were calculated for all RILs in both tests.
A visualization of these least square means and the standard error on
the estimation of these means for seed protein, seed oil, and seed yield
can be seen in figure 1. Both tests tended to have more RILs with high
protein rather than high oil when compared to the average of the checks,
and a moderate number had a yield average greater than the average of
the checks. Qualitatively, many RILs had a comparable yield to the
checks following the relatively large standard error associated with the
estimating the yield, and a qualitatively larger seed protein content
given the relatively small standard error associated with the estimation
of seed protein content. To separate means quantitatively, we performed
contrasts between each RIL mean and the average of the checks for each
test.

\hypertarget{yield-contrasts}{%
\subsection{Yield contrasts}\label{yield-contrasts}}

Several RILs in each of the tests had yield that was comparable to that
of their check cultivars as per contrasts between RIL means and the
average of the check cultivars for each test. No genotypes had yield
that was higher than that of the yield checks, however many had
comparable yield. Many of these genotypes with comparable yield also had
protein content that was superior to the check average in each test. No
genotypes had oil that was superior to the check cultivars, but some had
comparable yield and oil content as well as superior protein content.
Such genotypes are extremely valuable for production of soybean meal
with higher protein. As mentioned earlier, the meal protein of current
U.S. cultivars is mostly below the minimum market standard of 47.5\%.
The meal protein is calculated by a formula that takes both seed protein
and oil contents into account (reference from Pantalone). The meal
protein contents of these lines with high protein and normal oil
contents ranged from xx -- xx\% compared to the meal protein contents of
the checks (Supplementary table XX).

RILs were also evaluated for lodging and seed quality (data not shown?).
All RILs had seed quality that was on par with that of the high-yielding
checks however, however 11 genotypes had lodging scores that were
statistically significantly greater than the checks on the basis of
contrasts. These RILs were removed from consideration for recommendation
on the basis of their relatively poor agronomic performance

A detailed summary of all contrasts can be seen in supplementary table
1.

\hypertarget{genotypes-with-comparable-seed-yield-and-seed-oil-and-superior-seed-protein-content}{%
\subsection{Genotypes with comparable seed yield and seed oil and
superior seed protein
content}\label{genotypes-with-comparable-seed-yield-and-seed-oil-and-superior-seed-protein-content}}

Four genotypes had yield and oil that was similar to that of the yield
checks, and seed protein that was greater than the average of the
checks. These genotypes are N18-1635 from Test 1 and genotypes N18-1627,
N18-1643, and N18-1783 from Test 2. Summary data for these lines is
given in table 1.

\begin{table}[H]

\caption{Soybean genotypes with yield and seed oil comparable to check cultivars, and seed protein superior to check cultivars.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllrlllrlllrll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{4}{c}{Protein} & \multicolumn{4}{c}{Yield} & \multicolumn{4}{c}{Oil} \\
\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
Genotype & Test Name & Value & Rank & Test Average & Check Average & Value & Rank & Test Average & Check Average & Value & Rank & Test Average & Check Average\\
\midrule
N18-1635 & Jay Test 1 & 43.61 (106.5\%) & 16 & 45.09 & 40.94 & 46.45 (103.1\%) & 4 & 42.92 & 45.05 & 21.72 (96.7\%) & 3 & 20.89 & 22.45\\
\cmidrule{1-14}
N18-1783 &  & 46.74 (109.3\%) & 8 &  &  & 37.63 (86.6\%) & 19 &  &  & 20.63 (96.5\%) & 11 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
\cmidrule{11-12}
N18-1627 &  & 45.9 (107.3\%) & 9 &  &  & 46.01 (105.9\%) & 8 &  &  & 20.62 (96.5\%) & 12 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
\cmidrule{11-12}
N18-1643 & \multirow{-3}{*}{\raggedright\arraybackslash Jay Test 2} & 44.35 (103.7\%) & 13 & \multirow{-3}{*}{\raggedright\arraybackslash 45.7} & \multirow{-3}{*}{\raggedright\arraybackslash 42.77} & 40.54 (93.3\%) & 16 & \multirow{-3}{*}{\raggedright\arraybackslash 43.65} & \multirow{-3}{*}{\raggedright\arraybackslash 43.44} & 21.45 (100.4\%) & 7 & \multirow{-3}{*}{\raggedright\arraybackslash 20.7} & \multirow{-3}{*}{\raggedright\arraybackslash 21.37}\\
\bottomrule
\multicolumn{14}{l}{\rule{0pt}{1em}\textsuperscript{*} The genotype name.}\\
\multicolumn{14}{l}{\rule{0pt}{1em}\textsuperscript{\dag} The genotype marginal mean for the phenotype (value divided by check average).}\\
\multicolumn{14}{l}{\rule{0pt}{1em}\textsuperscript{\ddag} The ranking of this genotype for the phenotype within its test.}\\
\multicolumn{14}{l}{\rule{0pt}{1em}\textsuperscript{\S} The average phenotype value for all genotypes in the test.}\\
\multicolumn{14}{l}{\rule{0pt}{1em}\textsuperscript{\P} The average value of the checks in the test.}\\
\end{tabular}}
\end{table}

A strong negative correlation was observed in both populations between
seed oil and seed protein, and a weak inverse correlation was observed
between seed protein and seed yield. This observation matches well with
previous findings on the inverse correlation between seed protein and
seed oil, and between seed protein and seed yield. Many RILs had high
protein when compared with the check cultivars so it was not surprising
that relatively few had both a superior seed protein content, and a
comparable seed oil content.

These RILs can provide valuable germplasm that can be used to
simultaneously improve both seed oil and seed protein content without
compromising seed yield or agronomics. This can be of use both to
breeders who are looking to improve seed composition traits in soybean
and for growers who are seeking options for novel soybean varieties with
seed composition that is superior to that of elite cultivars.

\hypertarget{genotypes-with-comparable-yield-and-superior-protein}{%
\subsection{Genotypes with comparable yield and superior
protein}\label{genotypes-with-comparable-yield-and-superior-protein}}

Seventeen RILs had a similar yield and superior protein content to the
check cultivars. Summary data for the seed protein and seed yield of
these lines is given in table 2. From among these RILs, N18-1627 had the
highest seed yield relative to the checks included in the test with an
average yield that was 105.9\% that of the check average of Test 2.
Genotype N18-1763 had the highest seed protein content on average with a
protein content that was 116.8\% that of the check average of Test 1. We
observed a negative correlation between seed protein and seed yield in
both tests so RILs with seed protein content that was above the check
average tended to have a seed yield content that was comparatively
lower.

\begin{table}[H]

\caption{Soybean RILs with superior protein content and comparable yield performance to high-yielding check cultivars.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllrlllrll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{4}{c}{Protein} & \multicolumn{4}{c}{Yield} \\
\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10}
Genotype & Test Name & Value & Rank & Test Average & Check Average & Value & Rank & Test Average & Check Average\\
\midrule
N18-1763 &  & 47.82 (116.8\%) & 2 &  &  & 39.45 (87.6\%) & 17 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1855 &  & 47.52 (116.1\%) & 3 &  &  & 43.17 (95.8\%) & 10 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1595 &  & 47.36 (115.7\%) & 4 &  &  & 43.52 (96.6\%) & 9 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1632-1 &  & 46.64 (113.9\%) & 6 &  &  & 45.73 (101.5\%) & 6 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1620 &  & 43.63 (106.6\%) & 15 &  &  & 46.54 (103.3\%) & 3 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1635 &  & 43.61 (106.5\%) & 16 &  &  & 46.45 (103.1\%) & 4 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1731 &  & 46.1 (112.6\%) & 8 &  &  & 43.16 (95.8\%) & 11 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1674 &  & 45.19 (110.4\%) & 9 &  &  & 44.26 (98.2\%) & 7 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1682 &  & 45.11 (110.2\%) & 10 &  &  & 43.64 (96.9\%) & 8 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1572 &  & 45.08 (110.1\%) & 11 &  &  & 38.58 (85.6\%) & 19 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1751 & \multirow{-11}{*}{\raggedright\arraybackslash Jay Test 1} & 44.8 (109.4\%) & 12 & \multirow{-11}{*}{\raggedright\arraybackslash 45.09} & \multirow{-11}{*}{\raggedright\arraybackslash 40.94} & 41.59 (92.3\%) & 15 & \multirow{-11}{*}{\raggedright\arraybackslash 42.92} & \multirow{-11}{*}{\raggedright\arraybackslash 45.05}\\
\cmidrule{1-10}
N18-1761 &  & 49.1 (114.8\%) & 2 &  &  & 45.74 (105.3\%) & 9 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1575 &  & 48.67 (113.8\%) & 4 &  &  & 42.49 (97.8\%) & 12 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1769 &  & 48.58 (113.6\%) & 5 &  &  & 37.88 (87.2\%) & 18 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1783 &  & 46.74 (109.3\%) & 8 &  &  & 37.63 (86.6\%) & 19 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1627 &  & 45.9 (107.3\%) & 9 &  &  & 46.01 (105.9\%) & 8 &  & \\
\cmidrule{1-1}
\cmidrule{3-4}
\cmidrule{7-8}
N18-1643 & \multirow{-6}{*}{\raggedright\arraybackslash Jay Test 2} & 44.35 (103.7\%) & 13 & \multirow{-6}{*}{\raggedright\arraybackslash 45.7} & \multirow{-6}{*}{\raggedright\arraybackslash 42.77} & 40.54 (93.3\%) & 16 & \multirow{-6}{*}{\raggedright\arraybackslash 43.65} & \multirow{-6}{*}{\raggedright\arraybackslash 43.44}\\
\bottomrule
\multicolumn{10}{l}{\rule{0pt}{1em}\textsuperscript{*} The genotype name.}\\
\multicolumn{10}{l}{\rule{0pt}{1em}\textsuperscript{\dag} The genotype marginal mean for the phenotype (value divided by check average).}\\
\multicolumn{10}{l}{\rule{0pt}{1em}\textsuperscript{\ddag} The ranking of this genotype for the phenotype within its test.}\\
\multicolumn{10}{l}{\rule{0pt}{1em}\textsuperscript{\S} The average phenotype value for all genotypes in the test.}\\
\multicolumn{10}{l}{\rule{0pt}{1em}\textsuperscript{\P} The average value of the checks in the test.}\\
\end{tabular}}
\end{table}

However, among these genotypes are several with both average seed
protein and average seed yield that were greater than the averages of
the checks for each test. These N18-1632-1, N18-1620, N18-1635 from test
1 and RILs N18-1761 and N18-1627 from test 2. Of particular note are
genotypes N18-1632-1 from test 1 and N18-1761 from test 2. These two
RILs have a seed protein content that is substantially above that of the
average protein content of he checks while also maintaining an average
yield that is above the average yield of the checks. These RILs are
ideal candidates for use in breeding programs which seek to improve seed
protein content in particular without compromising seed yield or
argonomics.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

We have identified several genotypes with seed yield that is comparable
to existing cultivars that are commonly used in production. Many of
these genotypes have seed protein content that exceeds that of the check
cultivars, and some of these with superior protein have a seed oil
content that is comparable to these check cultivars as well. These
genotypes have good agronomic qualities as well and will provide both
breeders and growers with new options for genotypes that can be used in
production or used in breeding programs that seek to improve valuable
soybean seed composition traits with significantly reducing the yield.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %% optional
% \supplementary{The following are available online at www.mdpi.com/link, Figure S1: title, Table S1: title, Video S1: title.}
%
% % Only for the journal Methods and Protocols:
% % If you wish to submit a video article, please do so with any other supplementary material.
% % \supplementary{The following are available at www.mdpi.com/link: Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{All sources of funding of the study should be
disclosed. Please clearly indicate grants that you have received in
support of your research work. Clearly state if you received funds for
covering the costs to publish in open access.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short
paragraph specifying their individual contributions must be provided.
The following statements should be used ``X.X. and Y.Y. conceive and
designed the experiments; X.X. performed the experiments; X.X. and Y.Y.
analyzed the data; W.W. contributed reagents/materials/analysis tools;
Y.Y. wrote the paper.'\,' Authorship must be limited to those who have
contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{The authors declare no conflict of interest.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
RIL & Recombinant inbred line \\
RCBD & Randomized complete block design \\
CLA & Central crops research station \\
CAS & Caswell research farm \\
PLY & Tidewater research station \\
\end{tabular}}

\input{"appendix.tex"}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: internal bibliography
%=====================================
%\reftitle{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem[Author1(year)]{ref-journal}
%Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
%\bibitem[Author2(year)]{ref-book}
%Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
%\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\bibliography{mybibfile.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds \ldots\ldots{} are
available from the authors.}

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}